---
title: "Covariance estimation"
author: "Davy Leo"
date: "2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(igraph)
#install.packages("huge")
library(huge) # contain Glasso and many others estimations for covariance and precision matrix
# the package already includes MASS and igraph packages
#set.seed(41) # to have a good estimation at the end
library(shapes)#contains distcov to compute distance between covariance in an intrinsic way
```


We simulate some data.

```{r simulation}
p <- 10
smallw <- c(0.5,0,-1,0,0,0.4, rep(0, p-7))
W <- matrix(0,nrow=p,ncol=p)
W[row(W)>col(W)] <- smallw
W <- (W+t(W))
diag(W) <- 5
cat('matrix W :\n')
  round(W,2)
  cat('\n')
  cat(W)
n <- 10
x <- mvrnorm(n,rep(0,p), W)
```
We compute the sample covariance matrix of a single matrix, assuming mean = 0
```{r}
S <- (1/n)*t(x)%*%x
#S<-cov(x)
round(S,2)
cat(norm(S - W, "F"))

x <- mvrnorm(5,rep(0,p), W)
SLasso_res <- huge.glasso(x, lambda = sqrt(log(p)/n), cov.output=TRUE)
SLasso <- SLasso_res$icov[[1]]
round(SLasso,2)
```
We plot the observed error of the sample covariance and the lasso covariance estimate on a number of simulations
```{r}
##compute_errors() computes solutions with lasso and sample covariance to covariance estimation
#it returns the value of the error between the exact solution target and the computed solutions
#Errors function are LS/Lass(*) with Frobenius norm (F), spectral norm (2) and Riemannian distance (R)
#LS is the sample covariance and Lass the lasso solution
compute_errors <- function(target, n, index){
  p<-ncol(target)
  x <- mvrnorm(n, rep(0,p), target, empirical=TRUE)
  S <- (1/n)*t(x)%*%x
  SLasso_res <- huge.glasso(x, lambda = sqrt(log(p)/n), cov.output=TRUE)
  SLasso <- SLasso_res$icov[[1]]
#  Results$LSFrob[index] <- norm(S-target, "F")
#  Results$LSSpectral[index] <- norm(S-target, "2")
#  Results$LSRiem[index] <- distcov(S, target, "Riemannian")  
  
#  Results$LassFrob[index] <- norm(SLasso - target, "F")
#  Results$LassSpectral[index] <- norm(SLasso - target, "2")
#  Results$LassRiem[index] <- distcov(SLasso, target, "Riemannian")
  
  rbind(results, c(norm(S-target, "F"), norm(S-target, "2"),distcov(S, target, "Riemannian"), norm(SLasso - target, "F"), norm(SLasso - target, "2"), distcov(SLasso, target, "Riemannian") ))
  print(results)
}

Nsim <- 10
max_ratio <- 10
results <- data.frame(matrix(ncol = 6, nrow = Nsim))
colnames(results) <- c('SampCovFrob', 'SampCovSpec', 'SampCovRiem', 'LassCovFrob', 'LassCovSpec', 'LassCovRiem')
results_varying_ratio <- data.frame(matrix(ncol = 6, nrow = max_ratio))
colnames(results_varying_ratio) <- c('SampCovFrob', 'SampCovSpec', 'SampCovRiem', 'LassCovFrob', 'LassCovSpec', 'LassCovRiem')

target<-W
for(j in 1:max_ratio){
  n <- p*j 
  for(i in 1:Nsim){
    x <- mvrnorm(n, rep(0,p), target)
    S <- (1/n)*t(x)%*%x
    SLasso_res <- huge.glasso(x, lambda = sqrt(log(p)/n), cov.output=TRUE)
    SLasso <- SLasso_res$icov[[1]]
  #  Results$LSFrob[index] <- norm(S-target, "F")
  #  Results$LSSpectral[index] <- norm(S-target, "2")
  #  Results$LSRiem[index] <- distcov(S, target, "Riemannian")  
  #  Results$LassFrob[index] <- norm(SLasso - target, "F")
  #  Results$LassSpectral[index] <- norm(SLasso - target, "2")
  #  Results$LassRiem[index] <- distcov(SLasso, target, "Riemannian")
    
    results[i,] <- c(norm(S-target, "F"), norm(S-target, "2"),distcov(S, target, "Riemannian"), norm(SLasso - target, "F"), norm(SLasso - target, "2"), distcov(SLasso, target, "Riemannian"))
  }
  results_varying_ratio[j,] <- c(mean(results$SampCovFrob),mean(results$SampCovSpec),mean(results$SampCovRiem),mean(results$LassCovFrob),mean(results$LassCovSpec),mean(results$LassCovRiem))
}

```
```{r}
#jpeg("estimate_errors_low_dim.jpg")
ymax = max(results_varying_ratio)
ymin = min(results_varying_ratio)

plot(x=1:max_ratio, y=results_varying_ratio$SampCovFrob, ylim = c(ymin, ymax), ylab = "Error between estimate and target", xlab = "n/p", col="red", pch=1)
points(x=1:max_ratio, y=results_varying_ratio$LassCovFrob, col="blue", pch=1)

points(x=1:max_ratio, y=results_varying_ratio$SampCovSpec, col="red", pch=2)
points(x=1:max_ratio, y=results_varying_ratio$LassCovSpec, col="blue", pch=2)

points(x=1:max_ratio, y=results_varying_ratio$SampCovRiem, col="red", pch=3)
points(x=1:max_ratio, y=results_varying_ratio$LassCovRiem, col="blue", pch=3)

legend(x = max_ratio*0.9, y = 16, legend = c("Sample covariance", "Lasso", "Frob", "Spec", "Riem"), col = c("red", "blue", "black", "black", "black"), pch = c(10, 10, 1, 2, 3), ncol = 1)

```

```{r}
breaklist <-c(Nsim/10, 10)
breakmin <- max(breaklist, na.rm = FALSE)
hist(results$SampCovFrob,main=paste("Samp Frobenius norm"), breaks= breakmin)
hist(results$SampCovSpec,main=paste("Samp Spectral norm"), breaks=breakmin)
hist(results$SampCovRiem,main=paste("Samp Riem dist"), breaks=breakmin)
hist(results$LassCovFrob,main=paste("Lasso Frobenius norm"), breaks=breakmin)
hist(results$LassCovSpec,main=paste("Lasso spectral norm"), breaks=breakmin)
hist(results$LassCovRiem,main=paste("Lasso Riemannian distance"), breaks=breakmin)

```


We apply the Graphical Lasso procedure.

```{r high lambda}
lambda <- sqrt(log(p)/n)
res <- huge.glasso(x, lambda = 100*lambda)
What <- res$icov[[1]]

cat('True precision matrix:\n')
    round(W,2)
    cat('\n')
cat('Estimated precision matrix:\n')
    round(What,2)
    cat('\n')
```

We display the result.

```{r fisrt graphs}
A <- W
diag(A) <- 0
A <- (A!=0)
Ahat <- What
diag(Ahat) <- 0
Ahat <- (Ahat!=0)

par(mfrow=c(1,2))
g <- graph_from_adjacency_matrix(A,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="True graph")
g <- graph_from_adjacency_matrix(Ahat,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="Estimated graph")
```

This result was obtained with a given penalty weight $\lambda$. As we have a too sparse graph, we shall reduce $\lambda$.

```{r small lambda}
lambda <- 0.1*sqrt(log(p)/n)
res <- huge.glasso(x, lambda = lambda)
What <- res$icov[[1]]

A <- W
diag(A) <- 0
A <- (A!=0)
Ahat <- What
diag(Ahat) <- 0
Ahat <- (Ahat!=0)

par(mfrow=c(1,2))
g <- graph_from_adjacency_matrix(A,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="True graph")
g <- graph_from_adjacency_matrix(Ahat,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="Estimated graph")

```

We can use a data-driven choice of the parameter $\lambda$.

```{r select}
res <- huge(x, method='glasso', verbose=FALSE)
res.select <- huge.select(res, criterion='stars', verbose=FALSE) 
```
```{r, echo=FALSE }
cat('Final value of the lambda: ', res.select$opt.lambda, '\n')
cat(paste('Final value of the lambda: ', round(res.select$opt.lambda*sqrt(n/log(p)),3),'*sqrt(log(p)/n)'))
```


```{r final graphs}
A <- W
diag(A) <- 0
A <- (A!=0)
Ahat <- res.select$refit

par(mfrow=c(1,2))
g <- graph_from_adjacency_matrix(A,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="True graph")
g <- graph_from_adjacency_matrix(Ahat,mode='undirected')
plot(g, layout=layout.circle, edge.arrow.size=0.8,vertex.label.color = "black",
     vertex.label.cex=1,vertex.size=20, vertex.color="white", edge.color="black",
     main="Estimated graph")

```

It does not always work\dots